Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-03 overleaf/vectorization.tex
Made a pass.

\begin{itemize}
\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$. E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) is a same-level edge. 
%A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. E.g., 15 to 4 in Biometric is a same-level back-edge.
\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
%An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).

@@

There are two key phases of the algorithm. Phase 1 inserts raise dimension and drop dimension 
operations according to def-uses. E.g., if there is an inner-to-outer dependence, it inserts $\mathit{raise\_dim}$,
and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. After this phase operations work
on arrays of the corresponding dimensionality and we optimistically vectorize all arrays. For example, 
lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays.

@@

Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it).
%and Phase 4 explicitly turns operations into MOTION SIMD operations.

@@

We build a correctness argument that loosely follows Abstract Interpretation.
First we define the MPC Source syntax. The domain of MPC Source programs
expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
We then define the \emph{linearization} of an MPC Source program as an interpretation
over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
in terms of def-use relations. We define a partial order over elements of $C$ as well,
in terms of def-use relations in the concrete domain $C$. We state two theorems
that state (informally) that the schedule corresponding to the original program computes the same
result as the schedule corresponding to the vectorized program.

>>>>

\begin{itemize}
\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$. E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) is a same-level edge. 
%A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. E.g., 15 to 4 in Biometric is a same-level back-edge.
\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
\item inner-to-outer $X\rightarrow Y$ where $X$ is in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
%An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).

@@

There are two key phases of the algorithm. Phase 1 inserts raise dimension and drop dimension 
operations according to def-uses. E.g., if there is an inner-to-outer dependence, it inserts $\mathit{raise\_dim}$,
and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. 
For example, lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays. 
After this phase, operations work on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.

@@

Phases 3 reverses scalar expansion (this is an optional phase and our current implementation does not include it).
%and Phase 4 explicitly turns operations into MOTION SIMD operations.

@@

We build a correctness argument that loosely follows Abstract Interpretation.
First we define the MPC Source syntax. The domain of MPC Source programs
expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
We then define the \emph{linearization} of an MPC Source program as an interpretation
over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
in terms of def-use relations. We define a partial order over elements of $C$ as well,
in terms of def-use relations in the concrete domain $C$. The theorems state 
(informally) that the vectorized program preserves def-uses and thus computes the same 
result as the original program.
